### 2013-09-03 research/paper/Theory/mathematical_model.tex
corrections made Dr Kenny's partially done

 In the QSS  model $C_y$ is determined by an interpolating polynomial based on the stationary lift and drag data. The order of the interpolation polynomial has varied from study to study. For  example a $7^{th}$ polynomial order was used in \cite{Parkinson1964} and $3^{rd}$ order polynomial was used in \cite{Barrero-Gil2009}. \cite{Ng2005} concluded that using a $7^{th}$ order polynomial is sufficient and a polynomial higher than that of $7^{th}$ order polynomial neither provides a significantly better result nor does it exhibit an additional amplitudes of oscillation. Thus a $7 ^{th}$ order interpolating polynomial was incorporated in this present study.

@@

\cite{Joly2012} reported a reduction of displacement amplitude at low mass ratios. Therefore in order to account for it {\Tan what is "it" ??} a sinusoidal forcing function to the RHS of the oscillator model (Eq. \eqref{equationofmotion})was introduced, in order to represent forcing due to VIV. This method provided satisfactory results with the numerical simulations obtained at low mass ratios. In this study, the forcing due to VIV is incorporated using a sinusoidal forcing function $F_0\sin{\omega_{s}t}$ added to the RHS of the equation. $\omega_{s}$ and $F_0$ represents shedding frequency and the maximum force due to shedding respectively. Thus, the final equation is given by Eq. \eqref{final_equation_motion}.

@@

 The dissipated power due to the mechanical damping could be expressed as the harvested power output assuming that the other  power dissipation due to internal damping such as friction of the system is negligible. Therefore the mean power output could be given by Eq. \eqref{power}.

@@

The stationary data and the fluid-structure interaction (FSI) data were obtained using a higher order spectral element code which simulate 2D laminar flow. The Reynolds number was kept at 165 as it was pointed out by \cite{Sheard2009} and \cite{Tong2008} that the 3 dimensional transition for a square cylinder occurs at approximately Re=160. $F_0$ was kept at $0.4937$ which was obtained by using a simple linear interpolation on the data presented in \cite{Joly2012}. $\omega_s$ was set to $0.98$ winch was obtained by performing a power spectral analysis of the stationary data at $0^0$. Stationary $C_y$ data were obtained at different angles of attack ranging from $0^0$ to $16^0$. The average power was obtained by using Eq. \eqref{power} with data sets consisting substantial amount of peaks. Power data  at Re=22300 were obtained using input $C_y$ data in \cite{Parkinson1964} in order to provide a comparison between high and low Reynolds numbers. $m^*$ was kept at 1163 for \reynoldsnumber=22300 (Similar as \cite{Parkinson1964}) and $m^*=20$ for \reynoldsnumber=165. These parameters  were used throughout this study unless otherwise specified.

>>>>

 In the QSS  model $C_y$ is determined by an interpolating polynomial based on the lift and drag data  for flow over a stationary body at various angles of attack. The order of the interpolation polynomial has varied from study to study. For  example a $7^{th}$ polynomial order was used in \cite{Parkinson1964} and $3^{rd}$ order polynomial was used in \cite{Barrero-Gil2009}. \cite{Ng2005} concluded that using a $7^{th}$ order polynomial is sufficient and a polynomial higher than that of $7^{th}$ order polynomial neither provides a significantly better result nor does it exhibit an additional amplitudes of oscillation. Thus a $7 ^{th}$ order interpolating polynomial was incorporated in this present study.

@@

\cite{Joly2012} introduced  an addition of a sinusoidal forcing function to the hydrodynamic forcing. This enables to model to provide accurate predictions even at low mass ratios where galloping \hilight{....} suppressed or not present. In this study, the forcing due to VIV is incorporated using a sinusoidal forcing function $F_0\sin{\omega_{s}t}$ added to the RHS of the equation. $\omega_{s}$ and $F_0$ represents shedding frequency and the maximum force due to shedding respectively. Thus, the final equation is given by Eq. \eqref{final_equation_motion}.

@@

 The dissipated power due to the mechanical damping represents the ideal potential amount of harvested power output. Therefore the mean power output could be given by Eq. \eqref{power}.

@@

The stationary data and the fluid-structure interaction (FSI) data were obtained using a higher order spectral element code to simulate the two-dimensional laminar flow. The Reynolds number was kept at 165 as it was pointed out by \cite{Sheard2009} and \cite{Tong2008} that the 3 dimensional transition for a square cylinder occurs at approximately Re=160. $F_0$ was kept at $0.4937$ which was obtained by using a simple linear interpolation on the data presented in \cite{Joly2012}. Shedding frequency $\omega_s$ was set to $0.98$ winch was obtained by performing a power spectral analysis of the stationary data at $0^0$. Stationary $C_y$ data were obtained at different angles of attack ranging from $0^0$ to $16^0$. The average power was obtained by using Eq. \eqref{power} with data sets consisting substantial amount of peaks. Power data  at Re=22300 were obtained using input $C_y$ data in \cite{Parkinson1964} in order to provide a comparison between high and low Reynolds numbers. $m^*$ was kept at 1163 for \reynoldsnumber=22300 (Similar as \cite{Parkinson1964}) and $m^*=20$ for \reynoldsnumber=165. These parameters  were used throughout this study unless otherwise specified.

### 2013-09-03 research/paper/introduction/intro-1.0.tex
corrections made Dr Kenny's partially done

A simple structure that is susceptible to flow-induced vibrations that are suitable for energy extraction are slender structures,such as cylinders, elastically mounted perpendicular to a fluid stream. With regards to a slender body two common types of   flow induced vibrations are Vortex Induced Vibrations (VIV) and aeroelastic galloping. Significant research has been carried out by Bernitsas and his team on extracting useful energy from VIV. Some of their significant work includes investigating the influence  of physical parameters such as mass ratio (the ratio of the mass of the cylinder and the displaced fluid), Reynolds number, mechanical properties(\cite{Raghavan2010a} ,\cite{Lee2011b}) and location (effect of the bottom boundary) (\cite{Raghavan2009}). However,the possibility of extracting energy using aeroelastic galloping has not been thoroughly investigated. Some theoretical work was carried out by (\cite{Barrero-Gil2010a}). Utilizing galloping may be a more viable method to harness energy from flow induced vibrations as it is not bounded by a "lock-in" range of reduced velocities(ratio between the freestream velocity and the product of the natural frequency of the system and the characteristic length).Therefore it is preferable to investigate further the possibility of harnessing energy from flow induced vibrations using aeroelastic galloping.

Real life energy harvesting systems use high damping ratios where the energy generator (e.g electrical generator) puts a significant amount of damping into the system. Therefore it is crucial to investigate  the behaviour of aeroleastic galloping scenarios at high damping ratios in order to optimise the system to obtain a acceptable power output.Hence the focus of this paper is concentrated on investigating the mechanical power output of high-damped galloping systems in laminar flow.

According to \cite{Paidoussis2010},\cite{Glauert1919} has provided a criterion for galloping by considering the auto-rotation of an aerofoil.  \cite{DenHartog1956}  has provided a theoretical explanation for galloping for iced electric transmission lines. A non-linear theoretical aeroelastic model to predict the response of galloping was developed by \cite{Parkinson1964} based on the quasi-steady state (QSS) theory. Experimental lift and drag data on a fixed square prism at different angles of attack were used as an input for the theoretical model. It essentially used a curve fit of the transverse force to predict the galloping response. The study managed to achieve a good agreement with experimental (wind tunnel) data.\cite{Joly2012} have observed that finite element simulations shows a sudden change in amplitudes  below a critical values of the mass ratio, which the  (QSS) model fails to reproduce. The Parkinson's equation was essentially modified to account for the vortex shedding and managed to produce the effects to the amplitude at low mass ratios.
\cite{Barrero-Gil2010a} have investigated the possibility of extracting power from vibrations caused by galloping using quasi-steady state theory. In the conclusions of that paper it was pointed out that in order to obtain a high power to area ratio the mass-damping ($m^*\zeta$) parameter should be kept low as well as the frequency of oscillations should be carefully matched \hilight{have a good agreement with the size of the cross section}. Another interesting conclusion was that energy conversion systems which uses galloping could operate over  a large range of flow velocities unlike VIV energy harvesting systems where the factor of energy conversion has a strong dependence on the incoming flow velocity.

>>>>

An elastically mounted slender structure such as cylinders \hilight{......} which  are susceptible to flow-induced vibrations has the potential for energy extraction. With regards to a slender body two common types of   flow induced vibrations are Vortex Induced Vibrations (VIV) and aeroelastic galloping. Significant research has been carried out by Bernitsas and his team on extracting useful energy from VIV. Some of their significant work includes investigating the influence  of physical parameters such as mass ratio, Reynolds number, mechanical properties(\cite{Raghavan2010a} ,\cite{Lee2011b}) and location (effect of the bottom boundary) (\cite{Raghavan2009}). However,the possibility of extracting energy using aeroelastic galloping has not been thoroughly investigated. Some theoretical work was carried out by (\cite{Barrero-Gil2010a}). Utilizing galloping may be a more viable method to harness energy from flow induced vibrations as it is not bounded by a narrow "lock-in" range of reduced velocities(ratio between the freestream velocity and the product of the natural frequency of the system and the characteristic length). This study further explores the possibility of harnessing energy from flow induced vibrations using aeroelastic galloping.

Energy harvesting systems have high damping ratios because the energy generator (e.g electrical generator) puts a significant amount of damping into the system. Therefore it is crucial to investigate  the behaviour of aeroleastic galloping scenarios at high damping ratios in order to optimise the system to obtain a acceptable power output.Hence the focus of this paper is concentrated on investigating the mechanical power output of high-damped galloping systems.

According to \cite{Paidoussis2010},\cite{Glauert1919} has provided a criterion for galloping by considering the auto-rotation of an aerofoil.  \cite{DenHartog1956}  has provided a theoretical explanation for galloping for iced electric transmission lines. A non-linear theoretical aeroelastic model to predict the response of galloping was developed by \cite{Parkinson1964} based on the quasi-steady state (QSS) theory. Experimental lift and drag data on a fixed square prism at different angles of attack were used as an input for the theoretical model. It essentially used a curve fit of the transverse force to predict the galloping response. The study managed to achieve a good agreement with experimental data.\cite{Joly2012} have observed that finite element simulations shows a sudden change in amplitudes  below a critical values of the mass ratio, which the  (QSS) model fails to reproduce. The Parkinson's equation was essentially modified to account for the vortex shedding and managed to produce the effects to the amplitude at low mass ratios.
\cite{Barrero-Gil2010a} have investigated the possibility of extracting power from vibrations caused by galloping using quasi-steady state theory. In the conclusions of that paper it was pointed out that in order to obtain a high power to area ratio the mass-damping ($m^*\zeta$) parameter should be kept low as well as the frequency of oscillations should be carefully matched \hilight{have a good agreement with the size of the cross section}. Another interesting conclusion was that energy conversion systems which uses galloping could operate over  a large range of flow velocities unlike VIV energy harvesting systems where the factor of energy conversion has a strong dependence on the incoming flow velocity.
